DATA 621 Homework #3

### 2019-10-30 hw3/HW3.Rmd
Folded in Corey's work

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(corrplot)
library(caret)
library(rpart)
library(rpart.plot)
library(Amelia)
# Thank you Stack Overflow!
# A Prefix nulling hook.

@@

The following show the how predictors are distributed between the areas with crime rates higher than the median *blue) and below the median (red).  What we are looking for is variables that could split data into two groups.

@@

NOX seems to be the best variable to divide the data into the two groups

## Data Preparation
```{r}
ggplot(train, aes(rad , nox, color=as.factor(target))) +
  geom_point()
```

```{r}
library(FFTrees)

test <- test[names(train)]

fit <- FFTrees(target ~ ., train, test, main = "Crime Rate", decision.labels = c("Below Median", "Above Median"))

for (i in 1:6){
  p <- plot(fit, data = "test", tree=i)
  print(p)
}
```

```{r}
fit <- rpart(target ~ ., method="class", data=train)
rpart.plot(fit)
```

```{r}
get_high_nox <- function(x){
  if(x > 0.53){
    return("High")
  } else {
    return("Low")
  }
}

get_divider <- function(x, y){
  if(x < 10 & y < 0.6){
    return(1)
  } else {
    return(0)
  }
}

train <- train %>%
  rowwise() %>%
  mutate(high_nox = get_high_nox(nox),
         divider = get_divider(rad, nox))

test <- test %>%
  rowwise() %>%
  mutate(high_nox = get_high_nox(nox),
         divider = get_divider(rad, nox))
```

## Feature Engineering
```{r}

library(fuzzyjoin)

density_diff <- function(data,eval_data,var,target,new_column){

data[[paste0(var,"temp")]]<- data[[var]]

eval_data[[paste0(var,"temp")]]<- eval_data[[var]]

#standardize variable between 0 and 1
data[[var]] <- (data[[var]] - min(data[[var]]))/(max(data[[var]]) - min(data[[var]]))

eval_data[[var]] <- (eval_data[[var]] - min(eval_data[[var]]))/(max(eval_data[[var]]) - min(eval_data[[var]]))

@@

new_data <- merge(means,data, by.x ="std_var", by.y =var)
new_data$std_var <- NULL
new_data$join <- NULL
new_data[new_column]<- new_data$density_diff
new_data$density_diff <- NULL
###################same thing with eval data ############################

eval_data$join <- eval_data[[var]]
df2$join <- df2$x
temp <- difference_left_join(eval_data,df2, by="join", max_dist =.001)
means <- aggregate(ydiff ~ temp$join.x, temp, mean)

@@

eval_data2 <- merge(means,eval_data, by.x ="std_var", by.y =var)
eval_data2$std_var <- NULL
eval_data2$join <- NULL
eval_data2[new_column]<- eval_data2$density_diff
eval_data2$density_diff <- NULL

eval_data2[[var]] <- eval_data2[[paste0(var,"temp")]]
eval_data2[[paste0(var,"temp")]]<- NULL

new_data[[var]] <- new_data[[paste0(var,"temp")]]
new_data[[paste0(var,"temp")]]<- NULL

mylist <- list(g1,g3,new_data,eval_data2)
names(mylist)<- c("dist","dist_diff","new_data_training","new_data_eval")
return(mylist)

}

```

```{r}
baseline <- glm(target ~ ., family = binomial(link = "logit"), train)
summary(baseline)
test$baseline <- ifelse(predict(baseline, test) < 0.5, 0, 1)
test$baseline_yhat <- predict(baseline, test)
confusionMatrix(factor(test$baseline), factor(test$target))
```

#Creating Some New Variables

@@

str(train)
str(test)

```

@@

Applying occam's razor we will create a baseline that only has one predictor.  Any model will have to out preform this simplest model.

```{r}
baseline <- glm(target ~ nox, family = binomial(link = "logit"), train)
summary(baseline)
test$baseline <- ifelse(predict(baseline, test) < 0.5, 0, 1)
test$baseline_yhat <- predict(baseline, test)
confusionMatrix(factor(test$baseline), factor(test$target))
```

```{r}
model1 <- glm(target ~ high_nox + indus, family = binomial(link = "logit"), train)
summary(model1)
test$model1 <- ifelse(predict(model1, test) < 0.5, 0, 1)
test$model1_yhat <- predict(model1, test)
confusionMatrix(factor(test$model1), factor(test$target))
```

```{r}
model2 <- glm(target ~ divider, family = binomial(link = "logit"), train)
summary(model2)
test$model2 <- ifelse(predict(model2, test) < 0.5, 0, 1)
test$model2_yhat <- predict(model2, test)
confusionMatrix(factor(test$model2), factor(test$target))
```

```{r}

density_models <- glm(target ~ nox_density +indus_density+age_density+dis_density+rad_density+tax_density+ptratio_density+lstat_density+medv_density, family = binomial(link = "logit"), train)
summary(density_models)
test$density_models <- ifelse(predict(density_models, test) < 0.5, 0, 1)
test$density_models_yhat <- predict(density_models, test)
confusionMatrix(factor(test$density_models), factor(test$target))
```

@@

## pROC Output

```{r}
library(pROC)
par(pty = "s")
roc(test[["target"]], test[["density_models_yhat"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
par(pty = "m")

library(pROC)
par(pty = "s")
roc(test[["target"]], test[["model2_yhat"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
par(pty = "m")

library(pROC)
par(pty = "s")
roc(test[["target"]], test[["model1_yhat"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
par(pty = "m")
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA, message=FALSE, warning=FALSE)
library(tidyverse)
library(kableExtra)
library(corrplot)
library(caret)
library(Amelia)
library(pROC)
# Thank you Stack Overflow!
# A Prefix nulling hook.

@@

Next we will look at each potential predictor and how it is distributed across the target variable.

The following plots show how predictors are distributed between a positive target variable (areas with crime rates higher than the median, i.e. blue) and a negative target variable (areas with crime rates below the median, i.e. red). What we are looking for is variables that show way to split data into two groups.

@@

Looking at the plots above, the `nox` variable seems to be the best variable to divide the data into the two groups.

## Data Preparation

### Feature Engineering
```{r}
library(fuzzyjoin)
density_diff <- function(data,eval_data,var,target,new_column){
  data[[paste0(var,"temp")]] <- data[[var]]
  eval_data[[paste0(var,"temp")]] <- eval_data[[var]]

  #standardize variable between 0 and 1
  data[[var]] <- (data[[var]] - min(data[[var]]))/(max(data[[var]]) - min(data[[var]]))
  eval_data[[var]] <- (eval_data[[var]] - min(eval_data[[var]]))/(max(eval_data[[var]]) - min(eval_data[[var]]))

@@

  new_data <- merge(means,data, by.x ="std_var", by.y =var)
  new_data$std_var <- NULL
  new_data$join <- NULL
  new_data[new_column] <- new_data$density_diff
  new_data$density_diff <- NULL
  ###################same thing with eval data ############################
  eval_data$join <- eval_data[[var]]
  df2$join <- df2$x
  temp <- difference_left_join(eval_data,df2, by="join", max_dist =.001)
  means <- aggregate(ydiff ~ temp$join.x, temp, mean)

@@

  eval_data2 <- merge(means,eval_data, by.x ="std_var", by.y =var)
  eval_data2$std_var <- NULL
  eval_data2$join <- NULL
  eval_data2[new_column] <- eval_data2$density_diff
  eval_data2$density_diff <- NULL

  eval_data2[[var]] <- eval_data2[[paste0(var,"temp")]]
  eval_data2[[paste0(var,"temp")]] <- NULL

  new_data[[var]] <- new_data[[paste0(var,"temp")]]
  new_data[[paste0(var,"temp")]] <- NULL

  mylist <- list(g1,g3,new_data,eval_data2)
  names(mylist)<- c("dist","dist_diff","new_data_training","new_data_eval")
  return(mylist)
}
```

### Creating Some New Variables

@@

str(train)
str(test)
```

@@

We start by applying occam's razor and create a baseline model that only has one predictor. Any model we build beyond that will have to outperform this simplest model.

```{r}
baseline <- glm(target ~ nox, family = binomial(link = "logit"), train)
summary(baseline)
test$baseline <- ifelse(predict.glm(baseline, test,"response") >= 0.5,1,0)
cm <- confusionMatrix(factor(test$baseline), factor(test$target),"1")
results <- tibble(model = "baseline",predictors = 1,F1 = cm$byClass[7],
                  deviance=baseline$deviance, 
                  r2 = 1 - baseline$deviance/baseline$null.deviance,
                  aic=baseline$aic)
cm
```

Our baseline model is ok with an F1 score of `r cm$byClass[7]`.

Next we try adding every other variable, to build a full model. From here we can work backwards and eliminate non-significant predictors:

```{r}
fullmodel <- glm(target ~ ., family = binomial(link = "logit"), train)
summary(fullmodel)
test$fullmodel <- ifelse(predict.glm(fullmodel, test,"response") < 0.5, 0, 1)
cm <- confusionMatrix(factor(test$fullmodel), factor(test$target),"1")
results <- rbind(results,tibble(model = "fullmodel",
                                predictors = 12,F1 = cm$byClass[7],
                                deviance=fullmodel$deviance, 
                                r2 = 1-fullmodel$deviance/fullmodel$null.deviance,
                                aic=fullmodel$aic))
cm
```

The full model has an F1 score is `r cm$byClass[7]`, which is a bit higher than before. However, many variables do not seem to be significant.

After some backward elimination of non-significant predictor variables, we arrive at the following model:

```{r}
model1 <- glm(target ~ . -tax -rm -chas - age -zn -indus, 
              family = binomial(link = "logit"), 
              train)
summary(model1)
test$model1 <- ifelse(predict.glm(model1, test,"response") < 0.5, 0, 1)
cm <- confusionMatrix(factor(test$model1), factor(test$target),"1")
results <- rbind(results,tibble(model = "model1",
                                predictors = 6,F1 = cm$byClass[7],
                                deviance=model1$deviance, 
                                r2 = 1-model1$deviance/model1$null.deviance,
                                aic=model1$aic))
cm
```

```{r}
density_models <- glm(target ~ nox_density +indus_density+age_density+dis_density+rad_density+tax_density+ptratio_density+lstat_density+medv_density, family = binomial(link = "logit"), train)
summary(density_models)
test$density_models <- ifelse(predict(density_models, test) < 0.5, 0, 1)
test$density_models_yhat <- predict(density_models, test)
cm <- confusionMatrix(factor(test$density_models), factor(test$target))
results <- rbind(results,tibble(model = "density models",
                                predictors = 9,
                                F1 = cm$byClass[7],
                                deviance = density_models$deviance,
                                r2 = 1 - density_models$deviance / density_models$null.deviance,
                                aic = density_models$aic))
cm
```

Since the assignment mentions that the purpose is prediction, we will prefer F1 score as our measure of model success.

```{r}
kable(results)
```

Looking these three models together, `model1` looks like the best one. Although the full model scored an F1 that was ever-so-slightly higher on our test data set, `model1` has half the predictors and it's scores are almost exactly the same as `fullmodel`.

@@

### pROC Output

#### Baseline
```{r}
par(pty = "s")
roc(test[["target"]], test[["baseline"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
par(pty = "m")
```

#### Full Model
```{r}
par(pty = "s")
roc(test[["target"]], test[["fullmodel"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
par(pty = "m")
```

#### Model1
```{r}
par(pty = "s")
roc(test[["target"]], test[["model1"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
par(pty = "m")
```

#### Density Model
```{r}
par(pty = "s")
roc(test[["target"]], test[["density_models_yhat"]], plot = TRUE, legacy.axes = TRUE, print.auc = TRUE)
par(pty = "m")
```
